HealpixNet: Efficient spherical Convolutional Neural Network with Healpix sampling for cosmological applications

### 2018-07-18 scnn.tex
edits after meeting with Nathanaël

    \item (TK) make the bibliography display only a few name
    % \item (yes) Michael, do you agree with the title HealpixNet? I (Michaël) extensively modified the abstract to reflect that it is a special case of graph CNNs tailored to Healpix. Hence I now agree with the title. Feel free to edit. Ideally I'd like to have "graph" in the title (graph CNNs are quite hot these days), but I couldn't find how.
    % \item Emphasize the novelty of using it on Healpix in abstract, into and rest of the paper
    \item (TK) add info on comparison with histograms, so that it's just a demonstration
    % \item Finish section on Convolutions on graphs (Michael - Nathanael)
    % \item finish section on Efficient convolutions (Michael)
    % \item (done) Plot of the new filters from Mass maps
    % \item (done!, plot to be added) check the power spectrum after smoothing
    % \item (done!) upgrade the plot of pooling (Nathanael)
    \item (TK) possibly improve plot of pooling
	% \item (done!) new plot for network structure (teaser plot) (Michael) What do you think? Todo: link the points on the sphere to the fully connected \nati{I like the plot, but I am thinking about putting multiple spheres with different maps for the convolutiion instead…}
    \item (TK NP) if we have time make a new candidate for network plot (optional)
    % \item (done!) Figure 1 to be round
    \item (first version done) conclusions
    % \item (done!) new simulations with 3 arcmin smoothing
    % \item (done!) simulation with cross-validation for the histogram
    % \item (done!) Update result section according to the new simulations...
    % \item (done!) simulation PSD
    % \item (done!) Results figure, add sigma8
    % \item (more or less done!) possibly move the eigen decomposition and heat into appendix + add discussion (Nathanael I am not sure about the heat diffusion example...)
    \item Should the heat example go to appendix
    \item TK read the heat section again
    \item TK read the result section
    \item TK check the appendix
    \item TK optional - example of data (CMB (mask) + mass map + radio intensity)
    \item TK read everything
	\item MD should we keep those small appendices rather than merging them in the main text?
	\item MD add eigenvalues, filtering speed, irregular (random) sampling) to comparison with spherical harmonics
\end{itemize}

\begin{figure*}
	\centering
    % \includegraphics[width=\linewidth]{architecture}
	\includegraphics[width=\linewidth]{figure_architecture}
	\caption{Overall network architecture, showing here two convolutional layers acting as feature extractors followed by a fully connected layer with softmax acting as the classifier. A convolutional layer is based on five operations: convolution, non-linearity, batch normalization, down-sampling, and pooling. While most operations are agnostic to the data domain, the convolution and the down-sampling have to be adapted. In this paper, we propose first to model the sphere with a graph and to perform the convolution on the graph. Graphs are versatile data structures which can model any sampling, even irregular or partial. Second, we propose to exploit a hierarchical pixelization of the sphere for the pooling operation. It allows the network to analyze the data at multiple scales while preserving the spatial localization of features.
    \nati{Great plot! Could you make the sphere round? If you use the current code in paper plot notebook, they will be.}
    }
	\label{fig:architecture}
\end{figure*}

@@

We give a practical demonstration of the application of our package to cosmological model discrimination using convergence maps, similar to \citep{schmelze2017cosmologicalmodel}.
In a simplified scenario, we classify convergence maps on parts of a sphere into two cosmological models.
These models were designed to have the same power spectrum in range $\ell < 1000$.
We compare the performance of the spherical CNN to a baseline algorithm, an SVM classifier which takes pixel histograms or Power Spectrum Densities (PSD) of these maps as input.
The comparison is made as a function of the noise level on convergence maps and the area of the sphere used. Results show that our model is better at discriminating the maps, is more resilient to noise, and \todo{is faster?}.
\nati{I do not think we should argue for speed gain in this paper. We are probably much more efficient than scnns with the convolution done in the spectral domain. However compared to traditional methods, we are in most of the cases slower.}
\todo{I think we should, as that is our main argument against using the spherical harmonics. We are slower than 2D CNNs sure, but those are not adapted to the sphere. Or what do you mean by traditional methods?}

@@

The first approach, leveraging the continuous spherical convolution, has been proposed by~\citet{cohen2018spherical}. The goal of this contribution was to define a rotation-equivariant convolution on the sphere, i.e., a rotation of the input implies the same rotation in the output.
The resulting convolution is performed by (i) a spherical harmonic transform (SHT), i.e., a projection on the spherical harmonics, (ii) a  multiplication in the spectral domain, and (iii) an inverse SHT. Hence the computational cost of a convolution is essentially due to the two SHTs. A naive implementation of the SHT costs $\bO(N_{pix}^2)$ operations, where $N_{pix}$ is the number of pixels. There fortunately exists accelerated schemes for some sampling sets, see for example \citet{mohlenkamp1999fast} and \citet{rokhlin2006fast}.
The main advantage of this approach is that it provides a mathematically well defined rotation-equivariant network. Nevertheless, even with accelerated SHTs, the convolutions remain expensive, limiting the practical use of this approach.
For example with the HEALPix sampling, which was designed to have a fast SHT by being iso-latitude, the computational cost of the SHT is $\bO(N_{pix}^{3/2}) = \bO(N_{side}^3) = \bO(\ell_{max}^3)$, where $\ell_{max}$ is the largest order (angular frequency) \citep{gorski2005healpix}.\footnote{All pixels are placed on $N_{ring} = 4N_{side}-1 = \bO(\sqrt{N_{pix}})$ rings of constant latitude. Each ring has $\bO(\sqrt{N_{pix}})$ pixels. Thanks to this iso-latitude property, the SHT is computed using recurrence relations for Legendre polynomials on co-latitude and fast Fourier transforms (FFTs) on longitude. The computational cost is thus $\sqrt{N_{pix}}$ FFTs for a total cost of $\bO(N_{pix} \log(\sqrt{N_{pix}}))$, plus $\sqrt{N_{pix}}$ matrix multiplications of size $\sqrt{N_{pix}}$ for a total cost of $\bO(N_{pix}^{3/2})$.}
In comparison, filtering with our method scales as $\bO(N_{pix})$ (see \secref{efficient_convolution}). Moreover, \figref{filtering_speed} shows that a naive implementation of our method is more than ten times faster than the SHT implemented in HEALPix.
While a comparison is theoretically possible, the experiments
of~\cite{cohen2018spherical} are done using the geodesic grid sampling set and
do not provide any code for the HEALPix one. Hence, a practical comparison is
not possible. We also note that this approach cannot be trivially accelerated
if only one part of the sphere is used. This is impractical in the field of
cosmology as the data often observed only on some part of the sphere.
One could still pad the unused part of the sphere with zeros. However, this would significantly increase the computational cost of the method.
\nati{@mdeff: do you think we can say the method would be too slow for our dataset even without trying it? I am not so sure...}
\todo{slow: full sky maps are composed of millions of points. partial maps are hard to process}

A second direction has been followed in~\citet{boomsma2017spherical}. The idea is
to use the traditional 2-dimensional convolution on an irregular grid defined on the
sphere. In the paper, two grids are used: the equirectangular grid and the ``cubed
sphere'' grid defined by~\citet{ronchi1996cubed}. \nati{Do we need a plot of the equirectangular and cubed sphere sampling?} Because this
approach is based on the traditional 2-dimensional convolution, it is very likely to be the
most computationally efficient one. However, it suffers that it can only be used using very
specific grid sampling sets which are not including HEALPix. Furthermore, the
grid requirement makes it impossible for the convolution to capture the
spherical properties of the domain, i.e., the ``cubed sphere'' sampling is by
definition adapted to a cube and not to a sphere.
\todo{border effects? \nati{@mdeff what are you thinking about here?} I was thinking about irregularities in the sampling, i.e., the corners of the cube are much more sampled.}
% \nati{Michael: Under some very
% specific hypothesis, this second approach is a particular case of our method,
% i.e. 1) our method with a stupid sampling, 2) assumption that graph convolution
% on a grid == 2d convolution. Do you think we should mention that? I think we
% should not.}
% I think we should!

@@

The use of a graph to model the discretized sphere was considered by~\citet{khasanova2017graph} as well. The authors map omnidirectional images to the sphere, with the goal of correcting for distortions induced by the lens used for the acquisition.
Besides the different application covered by the paper, there are two important
differences. First the sampling is different as~\cite{khasanova2017graph} uses a
grid. Second, it uses a different graph CNN based on a different parameterization
of the convolution kernel.
As~\citep{boomsma2017spherical}, they use an equirectangular grid sampling. Moreover, they did not take advantage of a hierarchical pixelization of the sphere and resorted to dynamic pooling~\citep{kalchbrenner2014dcnn}.
Finally, while we do want our network to be rotation invariant, their statistical layer --- an operation which computes a set of spatially invariant statistics of the last feature map --- destroys a lot of information \todo{and hinders performance (should be empirically verified)}. While being completely invariant is a nice goal, CNNs for image recognition (arguably the most researched and developed architectures), are not built to be invariant to translation. As with the proposed spherical CNN, the fully connected layers at the end of the network depend on the spatial localization of features. While that may not be ideal, it was shown to work well in practice.

@@

The lowest possible resolution is given by the base partitioning of the surface into $N_{pix} = 12$ equal sized pixels (right-most sphere in \figref{pooling}). As each pixel is subdivided in four, the second coarser resolution is $N_{pix} = N_{side}^2 \cdot 12 = 2^2 \cdot 12 = 48$ pixels (middle sphere in \figref{pooling}), the third is $N_{pix} = N_{side}^2 \cdot 12 = 4^2 \cdot 12 = 192$ pixels, etc., where $N_{side} = 1, 2, 4, 8, \ldots$ is the grid resolution parameter.
High-resolutions HEALPix maps easily reach millions of pixels. \figref{example_maps} shows two examples.
\nati{The figure is beautiful, but should we really show it? If we target cosmologists, we can simply list some type of data. If we target a broader audience, maybe it makes sense to provide some illustration.}

@@

It turns out that using the graph construction described above, the graph Fourier modes resemble the spherical harmonics (see \figref{graph_harmonics}). That is a strong hint that the graph is able to capture the spherical properties of the HEALPix sampling. This topic is further discussed in \ref{sec:spherical_harmonics}.

\subsection{Convolution on graphs}

@@

Contrary to classical signal processing on Euclidean domains, the kernel $h$ has no single representation in the vertex domain and cannot be translated on the graph. It can however be localized on any vertex $v_i$ by the convolution with a Kronecker delta $\b \delta_i \in \R^{N_{pix}}$,\footnote{A Kronecker delta is the signal $\b \delta_i \in \R^{N_{pix}}$ that is zero everywhere except on vertex $v_i$ where it takes the value one.} that is $h(\L) \b \delta_i$, a column of $h(\L)$.
That is useful to visualize kernels.
If the graph is not regular, i.e., all nodes don't have the same number of neighbors, and all distances are not equal, the effect of the kernel will slightly differ from one node to another. While there is no perfect sampling of the sphere, these differences are negligible as the structure of the whole graph is very regular. However, when considering only parts of the sphere, one can observe important border effects (see \figref{border_effects} and~\ref{sec:border_effects}).
Finally, the graph convolution can be interpreted in the vertex domain as a scalar product with all localizations of the kernel:
\begin{equation*}
	(h(L) \f)_i = \langle h(L) \b \delta_i, \f \rangle.
\end{equation*}

@@

\begin{figure*}
	\centering
	\includegraphics[width=\linewidth]{gaussian_filters_sphere}
	\caption{Comparison of convolution with the graph and the spherical harmonics ($N_{side} = 16$).
	Top: Diffusion of a unit of heat for different times $t$ using the graph.
	Bottom: spherical symmetric Gaussian smoothing for different $\sigma$ (arcmin).
	Relative difference between graph convolution and spherical smoothing: $10.4$\%, $5.7$\%, $4.8$\%, $3.8$\% }
	\label{fig:gaussian_filters_comparizon}
\end{figure*}

@@

Following \citet{defferrard2016convolutional}, we define our filters as Chebyshev polynomials such as the convolution operation becomes
\begin{equation} \label{eqn:graph_convolution_cheby}
	h_\theta(\tL) \f = \sum_{k=0}^K \theta_k T_k(\tL) \f,
\end{equation}
where $\tL = \frac{2}{\lambda_{max}} \L - \b{I} = -\frac{2}{\lambda_{max}} \D^{-1/2} \W \D^{-1/2}$ is the rescaled Laplacian with eigenvalues in $[-1, 1]$ and $T_k(\cdot)$ is the Chebyshev polynomial of degree $k$, defined by the recursive relation $T_k(\tL) = 2\tL T_{k-1}(\tL) - T_{k-2}(\tL)$, $T_1(\tL) = \tL$, $T_0(\tL) = \b{I}$. While definitions \eqnref{graph_convolution_polynomial} and \eqnref{graph_convolution_cheby} both allow the representation of the same set of filters, Chebyshev polynomials are preferred as they form an orthogonal basis on the interval $[-1, 1]$. Moreover, they factor out $\tL^{k-2}$ from $T_k(\tL)$, e.g., the 3-neighborhood is not mixed with the 5-neighborhood. That results in easier learning when optimizing the parameters $\theta$.
Finally, the coefficients can be computed such as the truncated Chebyshev expansions gives an approximate minimax polynomial of a given filter \citep{hammond2011wavelets}. While that is irrelevant when learning filters, it is useful when smoothing a signal on the sphere by filtering it with an approximated low-pass filter.

@@

The main product of this contribution is HealpixNet,\footnote{\url{https://github.com/SwissDataScienceCenter/HealpixNet}} a small and easy-to-use python package for cosmologists to classify sky maps.
As a future work, we intend to 1) further investigate the link between the graph Fourier basis and the spherical harmonic and 2) compare HealpixNet to other spherical CNN using different sampling schemes.

@@

\section{Graph Fourier eigenvectors and spherical harmonics}
\label{sec:spherical_harmonics}

The eigenvalues of the graph Laplacian $\L$, shown in \figref{graph_eigenvalues}, are clearly organized in frequency groups of $2\ell + 1$ orders for each degree $\ell$.
The first 16 eigenvectors $[\b u_1, \ldots, \b u_{16}]$ of the graph graph Laplacian $\L$, forming the lower part of the graph Fourier basis $\U$, are shown in \figref{graph_harmonics}. While there is again a clear resemblance with the spherical harmonics, the small irregularities in the sampling (non-constant number of neighboring pixels and varying distance between pixels, see \figref{healpix_graph_4}) have an effect on the graph Fourier modes.
First, counter-intuitively, some eigenvectors will be localized~\citep{perraudin2018global}, i.e, they will span a small part of the sphere.
Second, as the resolution is increased with $N_{side} \rightarrow \infty$ and $N_{pix} \rightarrow \infty$, we are still unsure if the eigenvectors would converge towards the spherical harmonics. We believe that it might be the case for a different construction of the graph (fully connected) and that a proof could be built on top of the work of~\cite{belkin2007convergence}. That is however out of the scope of this contribution.

\begin{figure*}[!htb]
\centering
\includegraphics[width=\linewidth]{figures/eigenvectors}
\caption{The first 16 Fourier modes (eigenvectors of the graph Laplacian) of a graph constructed from the HEALPix sampling of the sphere. Eigenvectors 1--3 could be associated with spherical harmonics of degree $\ell=1$ and order $m=(-1,0,1)$, eigenvectors 4--8 with degree $\ell=2$ and order $m=(-2,-1,0,1,2)$, and eigenvectors 9--15 with degree $\ell=3$ and order $m=(-3,\ldots,3)$. Nevertheless, the graph eigenvectors are only approximating the real spherical harmonics.}
\label{fig:graph_harmonics}
\end{figure*}

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{figures/graph_eigenvalues}
	\caption{The eigenvalues $\bLambda$ of the graph Laplacian $\L = \U \bLambda \U\trans$, corresponding to squared frequencies, are clearly organized in groups. Each group corresponds to a degree $\ell$ of the spherical harmonics. Each degree has $2\ell+1$ modes. See also \figref{graph_harmonics}.}
	\label{fig:graph_eigenvalues}
\end{figure}

@@

\section{Border effects when only a part of the sphere is considered.}
\label{sec:border_effects}
The graph setting used through this contribution assumes reflective border conditions on a grid. While the complete sphere is not affected by this because it has no border, these border effects change slightly the effect of the convolution when working with only a part of the graph.  As depicted in \figref{border_effects}, the convolution of a Krnoneker near the border filters may no longer have a spherical profile.

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{figures/border_effects.pdf}
\caption{Graph convolution of a Kroneker placed at two different spots on the graph. The graph build from $1/12$ of the sphere and an Nside of 16. When the Kroneker is placed on the left corner, the resulting convolution is not circular anymore due to border effect. The graph setting assumes reflective border conditions.}
\label{fig:border_effects}
\end{figure}

>>>>

    \item (TK) make the bibliography display only a few name
    % \item (yes) Michael, do you agree with the title HealpixNet? I (Michaël) extensively modified the abstract to reflect that it is a special case of graph CNNs tailored to Healpix. Hence I now agree with the title. Feel free to edit. Ideally I'd like to have "graph" in the title (graph CNNs are quite hot these days), but I couldn't find how.
    % \item Emphasize the novelty of using it on Healpix in abstract, into and rest of the paper
    \item (TK) add info on comparison with histograms, so that it's just a demonstration
    % \item Finish section on Convolutions on graphs (Michael - Nathanael)
    % \item finish section on Efficient convolutions (Michael)
    % \item (done) Plot of the new filters from Mass maps
    % \item (done!, plot to be added) check the power spectrum after smoothing
    % \item (done!) upgrade the plot of pooling (Nathanael)
    \item (TK) possibly improve plot of pooling
	% \item (done!) new plot for network structure (teaser plot) (Michael) What do you think? Todo: link the points on the sphere to the fully connected \nati{I like the plot, but I am thinking about putting multiple spheres with different maps for the convolutiion instead…}
    \item (TK NP) if we have time make a new candidate for network plot (optional)
    % \item (done!) Figure 1 to be round
    \item (first version done) conclusions
    % \item (done!) new simulations with 3 arcmin smoothing
    % \item (done!) simulation with cross-validation for the histogram
    % \item (done!) Update result section according to the new simulations...
    % \item (done!) simulation PSD
    % \item (done!) Results figure, add sigma8
    % \item (more or less done!) possibly move the eigen decomposition and heat into appendix + add discussion (Nathanael I am not sure about the heat diffusion example...)
    \item Should the heat example go to appendix
    \item TK read the heat section again
    \item TK read the result section
    \item TK check the appendix
    \item TK optional - example of data (CMB (mask) + mass map + radio intensity)
    \item TK read everything
	\item MD add irregular (random) sampling to comparison with spherical harmonics? Could motivate people to think about different samplings.
	\item uniformize text size in figures
\end{itemize}

\begin{figure*}
	\centering
    % \includegraphics[width=\linewidth]{architecture}
	\includegraphics[width=\linewidth]{figure_architecture}
	\caption{Overall network architecture, showing here two convolutional layers acting as feature extractors followed by a fully connected layer with softmax acting as the classifier. A convolutional layer is based on five operations: convolution, non-linearity, batch normalization, down-sampling, and pooling. While most operations are agnostic to the data domain, the convolution and the down-sampling have to be adapted. In this paper, we propose first to model the sphere with a graph and to perform the convolution on the graph. Graphs are versatile data structures which can model any sampling, even irregular or partial. Second, we propose to exploit a hierarchical pixelization of the sphere for the pooling operation. It allows the network to analyze the data at multiple scales while preserving the spatial localization of features.}
	\label{fig:architecture}
\end{figure*}

@@

We give a practical demonstration of the application of our package to cosmological model discrimination using convergence maps, similar to \citep{schmelze2017cosmologicalmodel}.
In a simplified scenario, we classify convergence maps on parts of a sphere into two cosmological models.
These models were designed to have the same power spectrum in range $\ell < 1000$.
We compare the performance of our spherical CNN to a baseline algorithm, an SVM classifier which takes pixel histograms or power spectrum densities (PSDs) of these maps as input.
The comparison is made as a function of the additive noise level and the used area of the sphere. Results show that our model is better at discriminating the maps and is more resilient to noise.

@@

The first approach, leveraging the continuous spherical convolution, has been proposed by~\citet{cohen2018spherical}. The goal of this contribution was to define a rotation-equivariant convolution on the sphere, i.e., a rotation of the input implies the same rotation in the output.
The resulting convolution is performed by (i) a spherical harmonic transform (SHT), i.e., a projection on the spherical harmonics, (ii) a  multiplication in the spectral domain, and (iii) an inverse SHT. Hence the computational cost of a convolution is essentially due to the two SHTs. A naive implementation of the SHT costs $\bO(N_{pix}^2)$ operations, where $N_{pix}$ is the number of pixels. There fortunately exists accelerated schemes for some sampling sets, see for example \citet{mohlenkamp1999fast} and \citet{rokhlin2006fast}.
The main advantage of this approach is that it provides a mathematically well defined rotation-equivariant network. Nevertheless, even with accelerated SHTs, the convolutions remain expensive, limiting the practical use of this approach.
For example with the HEALPix sampling, which was designed to have a fast SHT by being iso-latitude, the computational cost of the SHT is $\bO(N_{pix}^{3/2}) = \bO(N_{side}^3) = \bO(\ell_{max}^3)$, where $\ell_{max}$ is the largest order (angular frequency) \citep{gorski2005healpix}.\footnote{All pixels are placed on $N_{ring} = 4N_{side}-1 = \bO(\sqrt{N_{pix}})$ rings of constant latitude. Each ring has $\bO(\sqrt{N_{pix}})$ pixels. Thanks to this iso-latitude property, the SHT is computed using recurrence relations for Legendre polynomials on co-latitude and fast Fourier transforms (FFTs) on longitude. The computational cost is thus $\sqrt{N_{pix}}$ FFTs for a total cost of $\bO(N_{pix} \log(\sqrt{N_{pix}}))$, plus $\sqrt{N_{pix}}$ matrix multiplications of size $\sqrt{N_{pix}}$ for a total cost of $\bO(N_{pix}^{3/2})$.}
In comparison, filtering with our method scales as $\bO(N_{pix})$ (see \secref{efficient_convolution}). Moreover, \figref{filtering_speed} shows that a naive implementation of our method is more than ten times faster than the SHT implemented in HEALPix.
While a comparison is theoretically possible, the experiments of~\cite{cohen2018spherical} are done using the geodesic grid sampling set and do not provide any code for the HEALPix one. Hence, a practical comparison is not possible.
We also note that this approach cannot be easily accelerated when data lies on a part of the sphere only. That is an important use case in cosmology as measurements are often partial, i.e., full-sky maps are rare. One could still fill the unseen part of the sphere with zeros, and avoid to compute empty integrals (on the latitude or longitude). It is however not straightforward to identify those empty directions, and computations would still be wasted (consider for example a ring which mostly contains zeros but a few measurements). With graphs however, one only pays for the used pixels. The cost is some distortions due to border effects (see \figref{border_effects} and \ref{sec:border_effects}). Those can however be mitigated by padding with zeros a small area around the measurements.

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{sphere_grids}
	\caption{Two realizations of a grid on the sphere. Left: the equirectangular grid, using equiangular spacing in a standard spherical-polar coordinate system. Right: An equiangular cubed-sphere grid, as described in \citet{ronchi1996cubed}. Figure from \citet{boomsma2017spherical}.}
	\label{fig:sphere_grids}
\end{figure}

A second direction has been followed in~\citet{boomsma2017spherical}. The idea is
to use the traditional 2-dimensional convolution on an irregular grid defined on the
sphere. In the paper, two grids are used: the equirectangular grid, and the cubed-sphere grid defined by~\citet{ronchi1996cubed}. See \figref{sphere_grids}. Because this
approach is based on the classic 2-dimensional convolution, it benefits from all the optimizations developed for images and is very likely to be 
the most computationally efficient one. However, it suffers that it can only be used using very
specific grid sampling sets which are not including HEALPix. Furthermore, the
grid requirement makes it impossible for the convolution to capture the
spherical properties of the domain, i.e., the cubed-sphere sampling is by
definition adapted to a cube and not to a sphere.
Those samplings are not ideal as their resolution is not constant across the sphere. Indeed, the singularities --- the poles for the equirectangular grid, the corners of the cube for the cubed-sphere --- are over-sampled. Equal areas of discrete elements is advantageous for physical measurements as white noise at the sampling frequency of the instrument gets integrated exactly into white noise in the pixel space, and sky signals are sampled without regional dependence.
% \nati{Michael: Under some very specific hypothesis, this second approach is a particular case of our method, i.e. (i) our method with a stupid sampling, (ii) assumption that graph convolution on a grid == 2d convolution. Do you think we should mention that? I think we should not.}

@@

The use of a graph to model the discretized sphere was considered by~\citet{khasanova2017graph} as well. The authors map omnidirectional images to the sphere, with the goal of correcting for distortions induced by the lens used for the acquisition.
Besides the different application covered by the paper, there are two important
differences. First the sampling is different as~\cite{khasanova2017graph} uses a
grid. Second, it uses a different graph CNN based on a different parameterization
of the convolution kernel.
As~\citep{boomsma2017spherical}, they use an equirectangular grid sampling. Moreover, they did not take advantage of a hierarchical pixelization of the sphere and resorted to dynamic pooling~\citep{kalchbrenner2014dcnn}.
Finally, while we do want our network to be rotation invariant, their statistical layer --- an operation which computes a set of spatially invariant statistics of the last feature map --- destroys a lot of information \todo{and hinders performance (should be empirically verified)}. Moreover, the chosen statistics have to be hand-crafted to capture statistics of interest. While developing completely invariant transforms is a  goal worth investigating, CNNs for image recognition (arguably the most researched and developed architectures), are not built to be fully invariant to translation. As with the proposed spherical CNN, the fully connected layers at the end of the network depend on the spatial localization of features. While that may not be ideal, it was shown to work well in practice.
\todo{The statistical layer, an idea that we reused in this work. See xxx.}

@@

The lowest possible resolution is given by the base partitioning of the surface into $N_{pix} = 12$ equal sized pixels (right-most sphere in \figref{pooling}). As each pixel is subdivided in four, the second coarser resolution is $N_{pix} = N_{side}^2 \cdot 12 = 2^2 \cdot 12 = 48$ pixels (middle sphere in \figref{pooling}), the third is $N_{pix} = N_{side}^2 \cdot 12 = 4^2 \cdot 12 = 192$ pixels, etc., where $N_{side} = 1, 2, 4, 8, \ldots$ is the grid resolution parameter.
High-resolutions HEALPix maps easily reach millions of pixels. \figref{example_maps} shows two examples.

@@

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{eigenvectors}
	\caption{The first 16 Fourier modes (eigenvectors of the graph Laplacian) of a graph constructed from the HEALPix sampling of the sphere. Eigenvectors 1--3 could be associated with spherical harmonics of degree $\ell=1$ and order $m=(-1,0,1)$, eigenvectors 4--8 with degree $\ell=2$ and order $m=(-2,-1,0,1,2)$, and eigenvectors 9--15 with degree $\ell=3$ and order $m=(-3,\ldots,3)$. Nevertheless, the graph eigenvectors are only approximating the real spherical harmonics.}
	\label{fig:graph_harmonics}
\end{figure}

It turns out that using the graph construction described above, the graph Fourier modes resemble the spherical harmonics (see \figref{graph_harmonics}). That is a strong hint that the graph is able to capture the spherical properties of the HEALPix sampling. This topic is further discussed in \ref{sec:comparison_spherical_harmonics}.

\subsection{Convolution on graphs}
\label{sec:graph_convolution}

@@

Contrary to classical signal processing on Euclidean domains, the kernel $h$ has no single representation in the vertex domain and cannot be translated on the graph. It can however be localized on any vertex $v_i$ by the convolution with a Kronecker delta $\b \delta_i \in \R^{N_{pix}}$,\footnote{A Kronecker delta is the signal $\b \delta_i \in \R^{N_{pix}}$ that is zero everywhere except on vertex $v_i$ where it takes the value one.} that is $h(\L) \b \delta_i$, a column of $h(\L)$.
That is useful to visualize kernels. \todo{nath: link with appendix}
If the graph is not regular, i.e., all nodes don't have the same number of neighbors, and all distances are not equal, the effect of the kernel will slightly differ from one node to another. While there is no perfect sampling of the sphere, these differences are negligible as the structure of the whole graph is very regular. However, when considering only parts of the sphere, one can observe important border effects (see \figref{border_effects} and~\ref{sec:border_effects}).
Finally, the graph convolution can be interpreted in the vertex domain as a scalar product with all localizations of the kernel:
\begin{equation} \label{eqn:graph_convolution_spatial}
	(h(L) \f)_i = \langle h(L) \b \delta_i, \f \rangle.
\end{equation}

@@

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{gaussian_filters_sphere}
	\caption{Comparison of convolution with the graph and the spherical harmonics ($N_{side} = 16$).
	Top: Diffusion of a unit of heat for different times $t$ using the graph.
	Bottom: spherical symmetric Gaussian smoothing for different $\sigma$ (arcmin).
	Relative difference between graph convolution and spherical smoothing: $10.4$\%, $5.7$\%, $4.8$\%, $3.8$\%.
	\todo{(i) Only three diffusion steps to have a bigger figure?}
\todo{(ii) The main issue with our filters is that their shape depends on the localization. It would be nice to measure the expectation and variance of the error over all locations. That would show how severe the distortion is.}}
	\label{fig:gaussian_filters_comparizon}
\end{figure}

@@

Following \citet{defferrard2016convolutional}, we define our filters as Chebyshev polynomials such as the convolution operation becomes
\begin{equation} \label{eqn:graph_convolution_cheby}
	h_\theta(\tL) \f = \sum_{k=0}^K \theta_k T_k(\tL) \f,
\end{equation}
where $\tL = \frac{2}{\lambda_{max}} \L - \b{I} = -\frac{2}{\lambda_{max}} \D^{-1/2} \W \D^{-1/2}$ is the rescaled Laplacian with eigenvalues in $[-1, 1]$ and $T_k(\cdot)$ is the Chebyshev polynomial of degree $k$, defined by the recursive relation $T_k(\tL) = 2\tL T_{k-1}(\tL) - T_{k-2}(\tL)$, $T_1(\tL) = \tL$, $T_0(\tL) = \b{I}$. While definitions \eqnref{graph_convolution_polynomial} and \eqnref{graph_convolution_cheby} both allow the representation of the same set of filters, Chebyshev polynomials are preferred as they form an orthogonal basis on the interval $[-1, 1]$. Moreover, they factor out $\tL^{k-2}$ from $T_k(\tL)$, e.g., the 3-neighborhood is not mixed with the 5-neighborhood. That results in easier learning when optimizing the parameters $\theta$.
Finally, the coefficients can be computed such as the truncated Chebyshev expansions gives an approximate minimax polynomial of a given filter \citep{hammond2011wavelets}. While that is irrelevant when learning filters, it is useful when smoothing a signal on the sphere by filtering it with an approximated low-pass filter.
\todo{(i) scale of $\theta_k$: they are mostly constant for Chebyshev polynomials. How are they for monomials?}
\todo{(ii) numerical stability of $\L^k \f$ v.s. $T_k(\L) \f$}
\todo{(iii) in the code, initialize $\theta_k$ with stdev adapted to $F_{in}$ and $K$}

@@

As shown in this contribution, modeling the discretized sphere, a 2D manifold in $\R^3$, by a graph enables an efficient convolution operation on any pixelization. Hence, it relaxes the iso-latitude constraint on sampling schemes who aim to enable fast convolutions. For applications which can tolerate small imprecisions in rotation invariance --- the fact that the action of a graph filter slightly depends on the location --- this relaxation might enable researchers to consider sampling sets with different properties.

The main product of this contribution is HealpixNet,\footnote{\url{https://github.com/SwissDataScienceCenter/HealpixNet}} a small and easy-to-use python package for cosmologists to classify sky maps.
As a future work, we intend to (i) further investigate the link between the graph Fourier basis and the spherical harmonic and (ii) compare HealpixNet to other spherical CNN using different sampling schemes.

@@

\section{Graph Fourier modes and spherical harmonics}
\label{sec:comparison_spherical_harmonics}
\label{sec:border_effects}

The eigenvalues of the graph Laplacian $\L$, shown in \figref{graph_eigenvalues}, are clearly organized in frequency groups of $2\ell + 1$ orders for each degree $\ell$.
The first 16 eigenvectors $[\b u_1, \ldots, \b u_{16}]$ of the graph Laplacian $\L$, forming the lower part of the graph Fourier basis $\U$, are shown in \figref{graph_harmonics}. \figref{subspace_harmonics_eigenvectors} shows the correspondence between the subspaces spanned by the graph Fourier modes and the spherical harmonics. While there is again a clear resemblance with the spherical harmonics, the small irregularities in the sampling (non-constant number of neighboring pixels and varying distance between pixels, see \figref{healpix_graph_4}) have an effect on the graph Fourier modes.
First, counter-intuitively, some eigenvectors will be localized~\citep{perraudin2018global}, i.e, they will span a small part of the sphere.
Second, as the resolution is increased with $N_{side} \rightarrow \infty$ and $N_{pix} \rightarrow \infty$, we are still unsure if the eigenvectors would converge towards the spherical harmonics. We believe that it might be the case for a different construction of the graph (fully connected) and that a proof could be built on top of the work of~\cite{belkin2007convergence}. That is however out of the scope of this contribution.

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{graph_eigenvalues}
	\caption{The eigenvalues $\bLambda$ of the graph Laplacian $\L = \U \bLambda \U\trans$, which corresponds to squared frequencies, are clearly organized in groups. Each group corresponds to a degree $\ell$ of the spherical harmonics. Each degree has $2\ell + 1$ orders. See also \figref{graph_harmonics}.}
	\label{fig:graph_eigenvalues}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{subspace_harmonics_eigenvectors}
	\caption{Correspondence between the subspaces spanned by the graph Fourier modes and the spherical harmonics. Each pixel on the plot is the dot product between one spherical harmonic and one graph Fourier mode $\b u_i$. The plot clearly shows how the subspaces are aligned: modes and harmonics which correspond to different degrees $\ell$ are mostly orthogonal to each other, while modes and harmonics which correspond to the same degree $\ell$ are linearly dependent. There is $2 \ell + 1$ orders per degree $\ell$.}
	\label{fig:subspace_harmonics_eigenvectors}
\end{figure}

The graph setting used through this contribution assumes reflective border conditions. That is irrelevant when working on the complete sphere as this surface has no border. When working with a part of the sphere however, its border slightly affects the convolution. As depicted in \figref{border_effects}, a filter localized near the border of a part of the sphere (via $h(\L) \b \delta_i$) no longer has an isotropic profile. These border effects can however be mitigated by padding with zeros a small area around the part of interest.

\begin{figure}[!ht]
	\centering
	\includegraphics[width=\linewidth]{border_effects}
	\caption{Convolution kernel (also called filter) $h$ localized in the center and left corner of a surface. A filter is localized on pixel $i$ as $h(\L) \b \delta_i$ (see \secref{graph_convolution} and \eqnref{graph_convolution_spatial}). The graph is built from $1/12^\text{th}$ of the sphere at $N_{side} = 16$. The filter is not isotropic anymore when localized on the left corner due to border effects. The graph representation of a manifold assumes reflective border conditions.}
	\label{fig:border_effects}
\end{figure}

@@
